### 2024-11-20 paper/background.tex
Split motivation into motiv&model

>>>>

\newcommand{\officialp}{official prose}
\newcommand{\spectecp}{SpecTec prose}

\red{TODO: explanation of the terms \officialp{} and \spectecp{}} \\
\red{TODO: remove redundant part in the fig} \\
\red{TODO: explanation of context(frame/label): in the background? or here?} \\
\red{TODO: the notion of meta-level interpreter in background} \\
\red{TODO: try finding better terminology rather than "model"} \\

\section{Control flow in official prose}
\label{sec:control flow in official prose}

% control flow structure in official prose
It might be easy to understand how the \officialp{} explains the control flow
of WebAssembly, if we assume that a WebAssembly code is loaded on a memory and
a pc points to the instruction to execute.
To describe control flow, it uses a structure named \textit{block} which
consitutes of a instruction sequence.
When executing the instructions in the block, pc can be changed to the starting
point of the block or the end of the block.

% an example of Wasm control flow in official prose
\begin{example}
\label{ex:br}
\begin{verbatim}
  // infinite loop
  (loop (result i32) (i32.const 42) (br 0) (unreachable) end) (unreachable)
\end{verbatim}
\end{example}

\cref{ex:br} is a WebAssembly code example that has \texttt{loop} and
\texttt{unreachable}
Here, result type of the \texttt{loop} is \texttt{i32} and the \texttt{loop}
has a block of three instructions: \texttt{i32.const}, \texttt{br}, and
\texttt{unreachable}.
\cref{fig:loop} is the \officialp{} of the \texttt{loop} instruction.
It says that the continuation is the start of the loop, the information is
stored in a label, and it \textit{enters} the block with the label.
The term \textit{enter} means that it pushes the label in the stack and makes
pc points to the first instruction in the block.
The \texttt{i32.const} instruction is executed first, which just pushes the
\texttt{i32} value \texttt{42} in the stack.
Then, The \texttt{br} instruction is executed.
\cref{fig:br} is the \officialp{} of the \texttt{br} instruction.
It pops the label from the stack, and makes pc points to the continuation of
the label, which is the start of the loop instruction.
As a result, the code example above is a infinite loop pushing \texttt{42}
forever without executing \texttt{unreachable} instructions.

\begin{figure}[h!]
    \centerline{\includegraphics[width=15cm]{fig/loop}}
    \caption[Enter the caption title here]{\texttt{loop} instruction} \label{fig:loop}
    \centerline{\includegraphics[width=15cm]{fig/br}}
    \caption[Enter the caption title here]{\texttt{br} instruction} \label{fig:br}
\end{figure}

% exiting label in official prose
There is also a special form of a behavior related to the block: an
\textit{exiting label}.
\cref{fig:exiting-label} is the \officialp{} of the \textit{exiting label}.
It is special because this behavior is performed without an explicit
WebAssembly instruction.
Rather, the behavior is performed when \textbf{the end of a block is reached}
without control instructions or runtime error.
The behavior is that the label is popped from the stack, and the pc changes to
the point after the end of the block.

\begin{figure}[h!]
    \centerline{\includegraphics[width=15cm]{fig/exiting}}
    \caption[Enter the caption title here]{exiting label} \label{fig:exiting-label}
\end{figure}

% an example of exiting label in official prose
\begin{example}
\label{ex:exit}
\begin{verbatim}
  // exiting label
  (loop (result i32) (i32.const 42) end) (f64.const 3.14)
\end{verbatim}
\end{example}

\cref{ex:exit} is a WebAssembly code example that has \texttt{loop} whose block is
only \texttt{i32.const}, and \texttt{f32.const}
When the \texttt{loop} instruction is executed, it enters the block with a label.
After \texttt{i32.const} is executed, the end of the block is reached.
As a result, exiting label occurs so that the label is popped from the stack,
pc changes to the point after the end of block: \texttt{f64.const}.
As a result, after executing the code, the value 42 and 3.14 pushed to the
stack.

% control flow structure with function call in official prose
Similar to the control flow using the block and the label, there is a function
call with a frame.
\cref{fig:invoke} is the \officialp{} of the function invocation which takes
place by a \texttt{call} instruction.
A frame is pushed to the stack and enters the block of the function body with a
label.
There is also a special behavior named returning from a function.
\cref{fig:returning} is the \officialp{} of the returning from a function.
When end of function is reached without without control instructions or runtime
error, The frame is popped from the stack, and the pc points to the next
instruction of the caller instruction.

\begin{figure}[h!]
    \centerline{\includegraphics[width=15cm]{fig/invoke}}
    \caption[Enter the caption title here]{function invocation} \label{fig:invoke}
    \centerline{\includegraphics[width=15cm]{fig/returning}}
    \caption[Enter the caption title here]{returning from a function} \label{fig:returning}
\end{figure}

### 2024-11-20 paper/model.tex
Split motivation into motiv&model

% !TEX root = main.tex

\chapter{Motivation}
\label{ch:motivation}
\noindent

\newcommand{\officialp}{official prose}
\newcommand{\spectecp}{SpecTec prose}

\red{TODO: explanation of the terms \officialp{} and \spectecp{}} \\
\red{TODO: remove redundant part in the fig} \\
\red{TODO: explanation of context(frame/label): in the background? or here?} \\
\red{TODO: the notion of meta-level interpreter in background} \\
\red{TODO: try finding better terminology rather than "model"} \\

\section{Control flow in official prose}
\label{sec:control flow in official prose}

% control flow structure in official prose
It might be easy to understand how the \officialp{} explains the control flow
of WebAssembly, if we assume that a WebAssembly code is loaded on a memory and
a pc points to the instruction to execute.
To describe control flow, it uses a structure named \textit{block} which
consitutes of a instruction sequence.
When executing the instructions in the block, pc can be changed to the starting
point of the block or the end of the block.

% an example of Wasm control flow in official prose
\begin{example}
\label{ex:br}
\begin{verbatim}
  // infinite loop
  (loop (result i32) (i32.const 42) (br 0) (unreachable) end) (unreachable)
\end{verbatim}
\end{example}

\cref{ex:br} is a WebAssembly code example that has \texttt{loop} and
\texttt{unreachable}
Here, result type of the \texttt{loop} is \texttt{i32} and the \texttt{loop}
has a block of three instructions: \texttt{i32.const}, \texttt{br}, and
\texttt{unreachable}.
\cref{fig:loop} is the \officialp{} of the \texttt{loop} instruction.
It says that the continuation is the start of the loop, the information is
stored in a label, and it \textit{enters} the block with the label.
The term \textit{enter} means that it pushes the label in the stack and makes
pc points to the first instruction in the block.
The \texttt{i32.const} instruction is executed first, which just pushes the
\texttt{i32} value \texttt{42} in the stack.
Then, The \texttt{br} instruction is executed.
\cref{fig:br} is the \officialp{} of the \texttt{br} instruction.
It pops the label from the stack, and makes pc points to the continuation of
the label, which is the start of the loop instruction.
As a result, the code example above is a infinite loop pushing \texttt{42}
forever without executing \texttt{unreachable} instructions.

\begin{figure}[h!]
    \centerline{\includegraphics[width=15cm]{fig/loop}}
    \caption[Enter the caption title here]{\texttt{loop} instruction} \label{fig:loop}
    \centerline{\includegraphics[width=15cm]{fig/br}}
    \caption[Enter the caption title here]{\texttt{br} instruction} \label{fig:br}
\end{figure}

% exiting label in official prose
There is also a special form of a behavior related to the block: an
\textit{exiting label}.
\cref{fig:exiting-label} is the \officialp{} of the \textit{exiting label}.
It is special because this behavior is performed without an explicit
WebAssembly instruction.
Rather, the behavior is performed when \textbf{the end of a block is reached}
without control instructions or runtime error.
The behavior is that the label is popped from the stack, and the pc changes to
the point after the end of the block.

\begin{figure}[h!]
    \centerline{\includegraphics[width=15cm]{fig/exiting}}
    \caption[Enter the caption title here]{exiting label} \label{fig:exiting-label}
\end{figure}

% an example of exiting label in official prose
\begin{example}
\label{ex:exit}
\begin{verbatim}
  // exiting label
  (loop (result i32) (i32.const 42) end) (f64.const 3.14)
\end{verbatim}
\end{example}

\cref{ex:exit} is a WebAssembly code example that has \texttt{loop} whose block is
only \texttt{i32.const}, and \texttt{f32.const}
When the \texttt{loop} instruction is executed, it enters the block with a label.
After \texttt{i32.const} is executed, the end of the block is reached.
As a result, exiting label occurs so that the label is popped from the stack,
pc changes to the point after the end of block: \texttt{f64.const}.
As a result, after executing the code, the value 42 and 3.14 pushed to the
stack.

% control flow structure with function call in official prose
Similar to the control flow using the block and the label, there is a function
call with a frame.
\cref{fig:invoke} is the \officialp{} of the function invocation which takes
place by a \texttt{call} instruction.
A frame is pushed to the stack and enters the block of the function body with a
label.
There is also a special behavior named returning from a function.
\cref{fig:returning} is the \officialp{} of the returning from a function.
When end of function is reached without without control instructions or runtime
error, The frame is popped from the stack, and the pc points to the next
instruction of the caller instruction.

\begin{figure}[h!]
    \centerline{\includegraphics[width=15cm]{fig/invoke}}
    \caption[Enter the caption title here]{function invocation} \label{fig:invoke}
    \centerline{\includegraphics[width=15cm]{fig/returning}}
    \caption[Enter the caption title here]{returning from a function} \label{fig:returning}
\end{figure}

>>>>

\chapter{Model}
\label{ch:Model}
\noindent
